Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/design.tex
Stream writeup

\subsection{Input streams and Transformations.}
\label{s:transforms}
In order to handle complicated input functions, Nail allows the programmer to
break out of Nail's generated code in a controlled fashion, by manipulating dependent fields and
streams. 
By default, Nail implements streams with in-memory buffers, but the programmer can override this implementation.
While parsing or generating a Nail grammar, there is always an implicit \emph{current stream} and
referred to as \cc{\$current} in the Nail language.

The programmer can define additional streams by writing a \emph{transformation}, a pair of
functions, one that generates new streams from existing streams during parsing and one that
combines streams during output generation. Temporary streams follow the naming and
scoping convention of dependent fields, but their names start with a \$-sign.

The parsing function of the transformation takes any number of streams, either temporary or the current,
and outputs any number of temporary streams. The parsing function may reposition and read from the
input streams and read the values of dependent fields, but not change their values. The generating
function has to be an inverse of the parsing function. It takes the same set of temporary streams
the parsing function returned and is supposed to change the values of the same set of values the
generating function produced.

>>>>

\subsection{Input streams and Transformations.}
\label{s:transforms}
Traditional parsers handle input one symbol at a time, beginning to end.
However, real-world formats often require non-linear parsing. Offset fields require a parser to move
to a different position in the input, possibly backwards. Size fields require the parser to stop
processing before the end of input has been reached and perhaps continue executing a parent parser.
Other cases, such as compressed data require more complicated processing on parts of the input
before it can be handled.

Nail introduces two concepts to handle this, \emph{streams} and \emph{transformations}. 
Streams represent a sequence of bytes that contain some external format. The parsers and generators
Nail generates always operate on an implicit stream named \cc{\$current} that they process front to
back, reading input or appending output.

Grammars can use the \cc{apply} combinator to parse or generate external data on a different stream,
inserting the result in the data model. Streams can be defined in two ways.

First, a rule can require streams as arguments, which can be passed in by the program that invokes
the generated parser.Every rule needs toe passed at least the implicit current stream.

Second, the programmer can create temporary streams in the grammars with \emph{transformations},
which are two arbitrary functions called during parsing and output generation.
The parsing function  takes any number of stream arguments and dependent field values
and produces any number of temporary streams. This function may reposition and read from the
input streams and read the values of dependent fields, but not change their contents and values. 
The generating function has to be an inverse of the parsing function. It takes the same number of temporary streams
the parsing function produces and and writes the same number of streams and dependent field
values the parsing function consumes.

Typically, the top level of most grammars is a rule that only takes a single stream, which may then
be broken up by various transformations and passed to sub-rules, which eventually parse various linear
fragment streams. Upon parsing, these fragment streams are generated and then combined by the
transforms.
